---
title: "Num_gen"
output: github_document
date: "2025-02-01"
---

```{r}
a = 2
c = 5
m = 5
seed = 1

ans = rep(0, 10)
ans[1] = (a * seed + c) %% m

```

```{r}
lcg = function(a, c, m, seed, nnum = 100) {
  ans = rep(0, nnum)
  ans[1] = (a * seed + c) %% m #the first number
  for (i in 2:nnum)
    ans[i] = (a * ans[i - 1] + c) %% m
  return(ans)
  
}

lcg(a = 10, c = 7, m = 1, seed = 4, nnum = 15)
```
The numbers Ui are supposed to follow a “discrete uniform” distribution on the integers 0 to m − 1. 

The period of a generator is the number of elements observed before the sequence begins to repeat itself.
The choice of a, c, m affects period and autocorrelation.

For the mixed LCG, the period will be m(full) if all the following conditions are satisfied:
1 a−1 is a multiple of q for every prime factor q of m.
2 if m is a multiple of 4, then a−1 is also a multiple of 4.
3 c is relatively prime (no common prime factors) to m.

simplified conditions
All these conditions for full period generator are met if
1. m = 2k, for some integer k.
2. a = 4b + 1, for some integer b.
3. c is an odd integer.
```{r}
randudat = lcg(a = 5, c = 7, m = 2^10, seed = 4, nnum = 500) / 2^10
par(mfrow = c(1,2))
plot(randudat)
hist(randudat)
```

## generate a rv from an arbitrary density f
```{r}
#Generate a vector of 10 exponential random numbers:
set.seed(2)
U = runif(10)
X = -log(1-U)

X
```
more examples
```{r}
#Generate a vector of 10 random variables with specified density
# f(x) = 2/x^3 for x > 1
n = 10
U = runif(n)
X = (1 - U)^(-0.5)

X

```
Suppose we are given a more complicated cdf:
0, x < 0,
x / 8 , 0 ≤ x < 2,
x^2/16, 2 ≤ x < 4,
1 , x ≥ 4.
```{r}
n = 10
U = runif(n)
X = (U < 1/4) * 8 * U + (U >= 1/4) * 4 * sqrt(U)

X
```
### generate discrete rvs
```{r}
# generate n bernoulli trials with probability p
easybinom = function(n, p) {
  out = sum(runif(n) < p)
  return(out)
}

easybinom(10, 0.7)
```
```{r}
#testing
runif(2*5)
mat = matrix(runif(2*5), nrow = 5) # need to generate all the numbers first, then arrange them in 5 rows...
res = mat < 0.5
res
res %*% rep(1 , 2)
as.vector(res %*% rep(1 , 2)) # the result is a 5x1 matrix, then convert to vector

```

```{r}
# function that returns the number of successes in several binomial trials
mybinom = function(num, n = 10, p = 0.5) {
  umat = matrix(runif(n * num), nrow = num) # each row contains n bernoulli trials
  res = umat < p
  return(as.vector(res %*% rep(1 , n)))
}
mybinom(num = 20)
```
## acceptance rejection method
use when can't obtain a closed form inverse cdf

```{r}
# fdens is a pre-defined target density function
# gdens is a pre-defined convenient density function
# M is the scaling factor
# x is the vector of random variables generated from gdens
accrej = function(fdens, gdens, M, x) {
  U = runif(length(x))
  selected = x[U <= (fdens(x) / (M * gdens(x)))]
  return(selected)
}
```
witch's hat
```{r}
unif02dens = function(y)
  return(1/2 * (y >= 0  & y <= 2))
witchshatdens = function(y)
  return((y >= 0 & y <= 1) * y + (y > 1 & y <= 2) * (2-y))
accrej = function(fdens, gdens, M, x) {
  U = runif(length(x))
  selected = x[U <= (fdens(x) / (M * gdens(x)))]
  return(selected)
}
x = runif(1000) * 2 #generating 1000 unif(0,2) rvs
z = accrej(fdens = witchshatdens, gdens = unif02dens, 2, x)
```
Histograms of candidate (left panel) and “accepted” (right panel) pseudorandom variables
```{r}
par(mfrow = c(1,2))
hist(x)
hist(z)
```

```{r}
ncandidates = 10
u = runif(ncandidates) #generate uniform random numbers from U(0,1)
x = tan(pi * (u - 0.5)) #generate candidate samp from a cauchy (gdens)

M = sqrt(2*pi/exp(1))
normaldens = dnorm(x)
cauchydens = dcauchy(x)

x[u <= normaldens/(M * cauchydens)]

```

